---
title: Archer Memes
author: Andrew Clark
date: '2017-10-26'
slug: archer-memes
categories:
  - packages
  - entertainment
tags:
  - slickR
  - meme
banner: 'banners/archerMeme.png'
description: ''
images: []
menu: ''
---

I have rather belatedly gotten around to viewing the adult animated series, [Archer](https://www.netflix.com/title/70171942), on Netflix. If you have not read up about the show, [wikipedia](https://en.wikipedia.org/wiki/Archer_(TV_series)) has good coverage and there is an in-depth viz project -[A data driven exploration of Archer](https://archervisualization.herokuapp.com/) - by [Mara Averick](https://twitter.com/dataandme) and [Elijah Meeks](https://twitter.com/Elijah_Meeks)


The subtitles often shout-out for a meme and I have clipped several, initially from Series 1 Ep 6. One way of viewing them is in a carousel such of that of the recently developed package, [htmlwidget slickR](https://github.com/metrumresearchgroup/slickR), from the Metrum research group

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE)
```

All we have to do is collect the relevant files and apply the slickR() function

I have 25 images in the carousel so it may take a time to process

```{r}
library(slickR)


  archerFiles <- list.files("img/archerMemes", full.names = TRUE, pattern = "*.JPG")

 slickR(obj=archerFiles, width= 650)
 
 
  
```

Just cycle through to view

Lots of them are standalone amusing, but the [meme package](https://github.com/GuangchuangYu/meme/) from
[Guangchuang Yu](https://twitter.com/guangchuangyu) makes it easy to add some upper text

```{r}

library(meme)

u <- "img/archerMemes/arousesMe.JPG"
meme(u, upper="The thought of a tidyverse tutorial",color="blue", size=2)


```

There is a meme_save() function so that you can swiftly add the memed image to a tweet

You can expect my tweets to be sprinkled with these from now on

***

Feel free to contact me if you have any questions or work offers

