---
title: EPL week 13
author: Andrew Clark
date: '2017-11-26'
slug: epl-week-13
categories:
  - eplWeekly
tags:
  - RcppRoll
banner: 'banners/wk13_2017.png'
description: 
images: []
menu: ''
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE, fig.align = 'center')
```



```{r setup}



library(tidyverse) #for data carpentry
library(plotly) # for plots
library(DT) # for tables
library(vembedr) # for videos
library(blogdown) # for shortcode

library(RcppRoll)

# library(stringr)
# library(vembedr) 
# library(blogdown)
# library(lubridate)
# library(crosstalk)
# library(forcats)
# library(glue)
# library(htmltools)

#library(sparkline)


standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")

# goals<- readRDS("data/soccer/goals.rds")
# playerGame<- readRDS("data/soccer/playerGame.rds")
# teamGames<- readRDS("data/soccer/teamGames.rds")
# assists<- readRDS("data/soccer/assists.rds")
#scoreLines<- readRDS("data/soccer/scoreLines.rds") #



```


## Match of the Day

The one top-six clash ended in stalemate leaving both Liverpool and Chelsea even further behind the Manchester clubs


```{r topGame}

embed_youtube("n_vQUh5X0KQ", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```

***


## Schedule difficulty

In spite of the season now more than a third gone, there remain some significant differences in terms of quality of opposition faced so far (based on finishing position last year)



```{r}
## last year's returning teams by position
old <- standings %>% 
  filter(season=="2016/17"&tmYrGameOrder==1) %>% 
  select(position=final_Pos,team) %>% 
  filter(position<18) 

# promoted clubs filling the bottom three spots - again by how they finished in Championship
new <- data.frame(team=c("Newcastle U","Brighton","Huddersfield"), position=c(18,19,20))

all <- bind_rows(old,new)


myFun <- function(x) {
  
  
  # need to remove selected team from the table and re-order
  change <- all %>% 
    filter(team!=x) %>% 
    mutate(order=row_number())
  
  # apply team's schedule to date
  standings %>% 
  filter(season=="2017/18"&team==x&tmYrGameOrder<=13) %>% 
  select(OppTeam) %>% 
  left_join(change,by=c("OppTeam"="team")) %>% 
  summarise(diff=mean(order)) %>% 
    pull(diff)
}

# map across allteasm
difficulty=map_dbl(all$team,myFun)

df <- data.frame(team=all$team,difficulty=difficulty)

df %>% 
  arrange(difficulty) %>% 
  #mutate(diff=row_number())%>%
  select(Most_difficult=team) %>% 
                         DT::datatable(class='compact stripe hover row-border order-column',rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE, pageLength = 20), width=300, height=700)


```

 Chelsea have had the hardest run to date and still have to face Swansea(H), Newcastle(H), West Ham(A), Huddersfield(A), Southampton(H) and Everton(A). You would have to think they would be disappointed with anything less than 15 points
 
Southampton's by contrast have Man City(A), Bournemouth(A), Arsenal(H), Leicester(H), Chelsea(A) and Huddersfield(H). The win over Everton will have given them a timely boost


***

<p class="factoid">In 29 starts with Sunderland last year, Jordan Pickford conceded an average of 1.72 goals per game. So far with Everton, it has been 2.15</p>

***

## So close for Burnley

Another controversial extra-time victory for Arsenal over Burnley has denied the Clarets from keeping four consecutive clean sheets in the EPL for the first time

They have to settle for equalling their best ever 

```{r}
burnley <- standings %>% 
  filter(team=="Burnley") %>% 
  arrange(gameDate)


counts = roll_sum(burnley$GA,n=4)

data.frame(goals=counts) %>% 
  mutate(order=row_number()) %>% 
  plot_ly(x=~order,y=~goals) %>% 
  add_lines() %>% 
  layout(title="Goals Conceded in 4 game-stretches - Burnley",
         xaxis=list(title="Game Order"),
         yaxis=list(title="Goals Against")) %>%  config(displayModeBar = F,showLink = F)
  


```



***

<p class="factoid">Only one team conceding more than an average of 2 goals a season has avoided relegation - Wigan 16th 2009/10. Other than Everton (see above). Stoke and West Ham are currently on the button </p>

***



***

## Tweet of the Week

```{r}  
shortcode("tweet", "933768157054955520") 
```



***


## Totally Football and Guardian Weekly Review Podcasts



<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6502679-pulis-p45-ed-narrative-busters-and-dulwich-hamlet-s-toilets-opposite-stand/embed/v4?eid=AQAAACyCE1oXOWMA" title="audioBoom player"></iframe>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6518587-shambolic-everton-manchester-city-s-marvels-and-dyche-s-deep-thinking-football-weekly/embed/v4?eid=AQAAAN1qHFo7d2MA" title="audioBoom player"></iframe>

***

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2017-11-21"&gameDate<="2017-11-27") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(team) %>% 
  select(Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=250)
```



```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2017-11-27") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```




## Final Factoid

<p class="factoid">Man City already look safe from relegation with 37 points. Only eight of the 66 teams relegated under the 38 game season with 38+ points have failed to survive</p>


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers
