---
title: EPL Week 19
author: Andrew Clark
date: '2017-12-26'
slug: epl-week-19
categories:
  - eplWeekly
tags: []
banner: 'banners/wk19_2017.png'
description: 
images: []
menu: ''
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE, fig.align = 'center')
```



```{r setup}



library(tidyverse) #for data carpentry
library(plotly) # for plots
library(DT) # for tables
library(vembedr) # for videos
library(blogdown) # for shortcode(tweet)







# library(stringr)
# library(vembedr) 
# library(blogdown)
 library(lubridate)
 library(crosstalk)
# library(forcats)
# library(glue)
 library(htmltools)
#library(RcppRoll) 

#library(sparkline)


standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")

goals<- readRDS("data/soccer/goals.rds")
playerGame<- readRDS("data/soccer/playerGame.rds")
teamGames<- readRDS("data/soccer/teamGames.rds")
# assists<- readRDS("data/soccer/assists.rds")
scoreLines<- readRDS("data/soccer/scoreLines.rds") #








```



## Match of the Day

Arsenal and Liverpool showing why they are fun to watch but won't be winning major trophies


```{r topGame}

embed_youtube("hzsCYuzgkao", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```


***
<p>


## Kane is Able

Some eyebrows were raised when Harry Kane was rated by a [Guardian poll](https://www.theguardian.com/football/ng-interactive/2017/dec/19/the-100-best-footballers-in-the-world-2017-interactive) of 169 experts  as the fifth best player in the world for 2017 but he certainly did his best to justify it with a hat-trick at high-flying Burnley

He now jointly - holds the record for most Premier League goals in a calendar year with one game remaining and is one of only three players to have led in that category and scored at the rate of more than a goal per 90 minutes played


```{r}

playerGame %>% 
    filter(PLAYERID!="OWNGOAL") %>% 
    select(PLAYERID,name,Gls,gameDate,mins) %>% 
    mutate(year=year(gameDate)) %>% 
    group_by(PLAYERID,name,year) %>% 
    summarize(tot=sum(Gls),totMins=sum(mins)) %>% 
    ungroup() %>% 
    arrange(desc(tot)) %>% 
    group_by(year) %>% 
    slice(1) %>% 
    mutate(gpg=round(90*tot/totMins,2)) %>% 
    plot_ly(x=~year,y=~tot) %>% 
 
    add_bars(color= ~year,
             hoverinfo="text",
             text=~paste0(name,
                          "<br>Goals: ",tot,
                          "<br>Gls per 90mins: ",gpg)) %>% 
 hide_colorbar() %>% 
  layout(
    title="Leading Scorer in Calendar year",
    xaxis=list(title=""),
    yaxis=list(title="Goals Scored")
  ) %>% 
  config(displayModeBar = F,showLink = F)
  
    

```

Kane was also the leader two years ago and will have Alan Shearer's four year record in his sights
Surprising winners? How about Robbie Keane 2007 and Wilfried Bony, 2014

***

Sort and filter the table, as desired


```{r}
 playerGame %>% 
    filter(PLAYERID!="OWNGOAL") %>% 
    select(PLAYERID,name,Gls,gameDate,mins,TEAMNAME) %>% 
    mutate(year=year(gameDate)) %>% 
    group_by(PLAYERID,name,year,TEAMNAME) %>% 
    summarize(tot=sum(Gls),totMins=sum(mins)) %>% 
    ungroup() %>% 
    arrange(desc(tot),totMins) %>% 
    group_by(year,TEAMNAME) %>%
    slice(1) %>% 
    mutate(gpg=round(90*tot/totMins,2)) %>% 
    arrange(desc(gpg)) %>% 
  ungroup() %>% 
  mutate(rank=row_number()) %>% 
  slice(2:11) %>% 
    select(rank,team=TEAMNAME,year,name,Goals=tot,`Goals per 90 mins`=gpg) %>%
                         DT::datatable(width=600,class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE))
  

```




***

<p class="factoid">Kane's 7th minute penalty ended a goal-drought three minutes shy of his season longest of 305 minutes </p>

## 


***








<p class="factoid">Last season, Raheem Sterling only scored three league goals after 24th September which was matched or bettered
by 108 other players. This season he currnetly is in sole position of 4th place</p>

***

## Tweet of the Week


```{r}  
shortcode("tweet", "944971910609113088") 
```



***

## Totally Football and Guardian Weekly Review Podcasts

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6564111-bristol-city-hit-the-big-time-clasico-chat-tom-ince-v-saido-berahino-and-the-doucoure-imperative/embed/v4?eid=AQAAAO7eQ1oPKWQA" title="audioBoom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6567565-christmas-carrolls-and-sid-s-ham-garage-football-weekly/embed/v4?eid=AQAAAMTbP1qNNmQA" title="audioBoom player"></iframe>

***

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2017-12-19"&gameDate<="2017-12-24") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(team) %>% 
  select(Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=250)
```

***
<p>

```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2017-12-24") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```


## Final Factoid

<p class="factoid"></p>


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers






