---
title: EPL Week 26
author: ''
date: '2018-02-10'
slug: epl-week-26
categories:
  - eplWeekly
tags: []
banner: 'banners/wk26_2017.png'
description: 
images: []
menu: ''

---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE,  fig.align = 'center')
```




```{r setup}


## Required

library(tidyverse) #for data carpentry
library(plotly) # for plots
library(DT) # for tables
library(vembedr) # for videos
library(blogdown) # for shortcode(tweet)







## Intermittent use

# library(vembedr) 
# library(blogdown)

# library(crosstalk)
# library(forcats)
# library(glue)
# library(htmltools)
#library(RcppRoll) 

#library(sparkline)


## Required data

standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")

## Intermittent use

managerGame<- readRDS("data/soccer/managerGame.rds")
#managers<- readRDS("data/soccer/managers.rds")
#goals<- readRDS("data/soccer/goals.rds")
#playerGame<- readRDS("data/soccer/playerGame.rds")
#playerClub<- readRDS("data/soccer/playerClub.rds")
#teamGames<- readRDS("data/soccer/teamGames.rds")
#teamCodes<- readRDS("data/soccer/teamCodes.rds")
# assists<- readRDS("data/soccer/assists.rds")
#scoreLines<- readRDS("data/soccer/scoreLines.rds") #



```

__For the remainder of the season, I will be travelling with a back up laptop so please excuse any shortfall in posts__

***

This week's crisis team, Chelsea.  


## Match of the Day



```{r topGame}

embed_youtube("MT6riYHiNUQ", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```

***
<p class="factoid">Jordon Ibe has picked up 3 assists in his most recent two games. He managed just one in 25 appearances for Bournemouth last season</p>

***

## Win Percentages

in spite of spending well in excess of 200 million pounds on players since last season's title-winning season, Chelsea are in some danger of failing to make the top 4 and the end of Conte's reign as manager now appears to be a matter of when rather than if

He and Guardiola have now managed 64 games in the Premier League with each winning 45

Here are how the top manager/team combinations all-time(min 6 games) have started their incumbencies

```{r }

top5 <-managerGame %>% 
  mutate(cat=ifelse(res=="Win",1,0)) %>% 
  select(name,TEAMNAME,res,cat) %>% 
  group_by(name,TEAMNAME) %>% 
  summarize(games=n(),wins=sum(cat), win_pc=round(100*wins/games,2)) %>% 
  filter(games>63) %>% 
  arrange(desc(win_pc)) %>% 
  head(5)

test <- top5 %>% 
  left_join(managerGame) %>% 
  select(name,TEAMNAME,res,gameDate) %>% 
  mutate(cat=ifelse(res=="Win",1,0)) %>% 
  group_by(name,TEAMNAME) %>% 
  arrange(gameDate) %>% 
  mutate(gameOrder=row_number(),cumwins=cumsum(cat), win_pc=round(100*cumwins/gameOrder,2)) 

library(glue)

test %>% 
  filter(gameOrder<65) %>% 
  mutate(manTeam=glue("{name} ({TEAMNAME})")) %>% 
  plot_ly(x=~gameOrder,y=~win_pc, color=~manTeam) %>% 
  add_lines() %>% 
  layout(title="Leading Managers Cumulative Win %",
         xaxis=list(title="Games Managed"),
         yaxis=list(title="% Wins")) %>%  config(displayModeBar = F,showLink = F)



```

At the 20 game mark, Conte had an 80% win rate but this has been on the slide in contrast to Guardiola's upward trend - particularly marked at the beginning of this season.
  However, Mourinho's 78% trumps both over the first seaon and two-thirds in the Premier League 




***

<p class="factoid">Benteke may be struggling to score, but at least he has 7 assists - with 4 the most he has recorded in any previous season</p>
***

## Consecutive thumpings

A late collapse to Watford consigned Chelsea to consecutive losses by three goal margins for the first time since they were beaten 1-4 by Man Utd. and 3-0 by Blackburn, two powerhouses in October 1995

Here are the most recent occurrences for the 38 teams that have suffered the indignation at least once

```{r}

standings %>% 
  select(team,gameDate,OppTeam,GF,GA) %>% 
  group_by(team) %>% 
  arrange(desc(gameDate)) %>% 
  mutate(cat=ifelse(GA-GF>2,1,0),leadCat=lead(cat),consec=cat+leadCat) %>% 
  filter(consec==2) %>% 
  group_by(team) %>%
  slice(1) %>% 
  arrange(desc(gameDate)) %>% 
  select(date=gameDate,team,GF,GA,opponents=OppTeam)%>%
  DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE))
         
```
<p>

Six teams have already had consecutive drubbings handed to them this season but Premier League ever-presents, Arsenal and Liverpool, have never suffered the fate


***

## Tweet of the Week 



```{r}  
shortcode("tweet", "962020532836167680") 
```


## Totally Football and Guardian Weekly Review Podcasts

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6641237-karma-mystery-tours-and-the-lepers-v-the-scoundrels/embed/v4?eid=AQAAAOCRfVpVVmUA" title="Audioboom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6641577-magical-mo-salah-marvellous-misses-and-celebrity-barracking-football-weekly/embed/v4?eid=AQAAAJuRfVqpV2UA" title="Audioboom player"></iframe>

***

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2018-02-02" &gameDate<="2018-02-06") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(gameDate,team) %>% 
  select(Date=gameDate,Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=400)
```

***
<p>

```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2018-02-06") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```


## Final Factoid

<p class="factoid">Looking like a 3-way race for Player of Season with pre-season favourit de Bruyne(7 goals, 22 assists) being challenged by Salah(21/11) and Kane(22/8)</p>


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers








