---
title: EPL Week 23
author: Andrew Clark
date: '2018-01-16'
slug: epl-week-23
categories:
  - eplWeekly
tags: []
banner: 'banners/wk23_2017.png'
description: 
images: []
menu: ''

---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE,  fig.align = 'center')
```



```{r setup}


## Required

library(tidyverse) #for data carpentry
library(plotly) # for plots
library(DT) # for tables
library(vembedr) # for videos
library(blogdown) # for shortcode(tweet)







## Intermittent use

# library(vembedr) 
# library(blogdown)

# library(crosstalk)
# library(forcats)
# library(glue)
# library(htmltools)
#library(RcppRoll) 

#library(sparkline)


## Required data

standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")

## Intermittent use

#managers<- readRDS("data/soccer/managers.rds")
#goals<- readRDS("data/soccer/goals.rds")
playerGame<- readRDS("data/soccer/playerGame.rds")
#playerClub<- readRDS("data/soccer/playerClub.rds")
#teamGames<- readRDS("data/soccer/teamGames.rds")
#teamCodes<- readRDS("data/soccer/teamCodes.rds")
# assists<- readRDS("data/soccer/assists.rds")
#scoreLines<- readRDS("data/soccer/scoreLines.rds") #








```

Man City's unbeaten run ended by a Coutinho-less Liverpool in seven goal thriller


## Match of the Day



```{r topGame}

embed_youtube("S326V6ytmts", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```


***


## Subs

On Saturday an oddity occurred. The kind of 'Dog didn't bark in the night' sort. Roy Hodgson, with his team hanging on to a precious three points for what would make a 'massive win" failed to make a substitution: a norm for time-wasting

He may have been giving a message to the hierarchy that with the club's injury problems he did not trust his subs to be up to it and they need to make purchases in the transfer window. Perhaps he was so engrossed in the game - he is 70, after all.

Whatever the reason, it was the first time it has happened for any team in the Premier League this season

See how this has changed over time.


```{r sub_use_by_year}

 playerGame %>% 
   filter(mins>0) %>% 
   group_by(TEAMMATCHID,season,TEAMNAME,gameDate) %>% 
   tally() %>% 
   arrange(gameDate) %>% 
   mutate(subs=n-11) %>% 
   group_by(season,subs) %>% 
   tally() %>% 
   ungroup() %>% 
   group_by(season) %>% 
   mutate(pc=100*nn/sum(nn)) %>% 
   plot_ly(x=~season,y=~pc) %>% 
   group_by(subs) %>% 
   add_bars(x=~season,y=~pc, color=~as.factor(subs)) %>% 
   layout(title="Number of Subs used by Match by Team",
     yaxis = list(title = 'Per Cent'),
          xaxis = list(title=""),
          margin=list(b=80),
          barmode = 'stack') %>%  config(displayModeBar = F,showLink = F) %>%  config(displayModeBar = F,showLink = F)
          
          

```
 
 In 2015/16 there were also no occurrences though it picked up a bit last year - when three substitutes were used a record 82.6%  of the time
 
 ***

The rules have changed over the years

* From the 1992-93 season, teams could name three substitutes, one of whom had to be a goalkeeper, and could use any two of them for any reason
* From the 1994-95 season, teams could name and use up to three substitutes (one of whom could only be used to replace the goalkeeper) for any reason
* From the 1995-96 season, teams could name and use up to three substitutes for any reason with no restriction on positions
* From the 1996-97 season, teams could name up to five and use up to three substitutes for any reason
* From the 2008-09 season, teams could name up to seven and use up to three substitutes for any reason

(Thanks Wikipedia)

***

<p class="factoid">With Stoke replacing Mark Hughes with Paul Lambert, there have been seven changes since the season started and, outside of the top seven teams, only Eddie Howe, at Bournemouth, was a Premier League manager at the end of last season</p>
***

## Recovery from Doom

Arguably, particularly given their long injury-list, the best performance has been by Roy Hodgson at Crystal Palace. After four games of Frank de Boer and three more defeats including away fixtures at both Manchester clubs, Palace had recorded the worst ever seven game start to a Premier League season with no points, no goals scored and 17 conceded

How have the next nine worst-starting clubs fared between the 7th and 23rd rounds and on until the final standings


```{r slow_start}


worst <- standings %>% 
  filter(tmYrGameOrder==7) %>% 
  arrange(cumPts,cumGD) %>% 
  slice(1:10) %>% 
  select(team,season,cumPts,cumGD,final_Pos,position) %>% 
  mutate(games=7)  
  

at23 <- standings %>% 
  filter(tmYrGameOrder==23) %>% 
  right_join(worst,by=c("team","season")) %>% 
  select(team,season,cumPts=cumPts.x,final_Pos=final_Pos.x,position=position.x) %>% 
  mutate(games=23)


final <- standings %>% 
  filter(tmYrGameOrder==38) %>% 
  right_join(worst,by=c("team","season")) %>% 
  select(team,season,cumPts=cumPts.x,final_Pos=final_Pos.x,position=position.x) %>% 
  mutate(games=38)


df <- bind_rows(worst,at23,final)

df %>% 
  group_by(team) %>% 
  mutate(combo=paste0(team," ",season)) %>% 
  plot_ly(x=~games,y=~cumPts) %>% 
  #add_lines(color=I("lightgrey")) %>% 
  add_markers(x=~games,y=~jitter(cumPts),color=~combo,
              hoverinfo="text",
              text=~paste0(combo,
                           "<br>Points: ",cumPts,
                           "<br>Position: ",position,
                           "<br>Finished: ",final_Pos)) %>% 
  layout(title="Progress of 10 Teams with Worst 7 Game Start",
         xaxis=list(title="Games Played"),
         yaxis=list(title="Cumulative Points")) %>%  config(displayModeBar = F,showLink = F) %>%  config(displayModeBar = F,showLink = F)
         
```

In spite of Palace's achievements, the Newcastle team featuring Alan Shearer (23 goals), Nolberto Solano(20 assists) and Gary Speed fared even better after their sticky start.    
 They were already six points clear of relegation after 23 games and finished 11th - only six points away from a top 6 place.      
 
 Currently only two points off 9th-placed-Everton, a top half finish is certainly feasible for Hodgson's team - though even 17th spot will be acceptable given the initial circumstances



***
<p class="factoid">Stoke have already conceded 50 goals this season, eight more than any other team. Only twice, has the team conceding the most goals avoided relegation. And both Middlesbrough (1996/97) and Fulham(2006/7) only let in a total of 60. Stoke are on pace for 83</p>
***



***

## Tweet of the Week



```{r}  
shortcode("tweet", "952662805768847365") 
```



***

## Totally Football and Guardian Weekly Review Podcasts

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6598399-klopp-clips-city-s-wings-an-apology-to-david-moyes-and-the-greatest-asian-player-of-all-time/embed/v4?eid=AQAAALDqXFr_rmQA" title="Audioboom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6598530-liverpool-spoil-manchester-city-s-party-while-stoke-and-wales-have-new-arrivals-football-weekly/embed/v4?eid=AQAAAOD8XFqCr2QA" title="Audioboom player"></iframe>

***

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2018-01-05" &gameDate<="2018-01-16") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(gameDate,team) %>% 
  select(Date=gameDate,Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=400)
```

***
<p>

```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2018-01-16") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```


## Final Factoid

<p class="factoid">Harry Kane now tops the goal charts for Spurs in the Premier League. Two more will see him reach the century mark, becoming just the 15th player to reach this level for one club</p>


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers





