---
title: EPL Week 24
author: ''
date: '2018-01-25'
slug: epl-week-24
categories:
  - eplWeekly
tags: []
banner: 'banners/wk24_2017.png'
description: 
images: []
menu: ''

---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE,  fig.align = 'center')
```



```{r setup}


## Required

library(tidyverse) #for data carpentry
library(plotly) # for plots
library(DT) # for tables
library(vembedr) # for videos
library(blogdown) # for shortcode(tweet)







## Intermittent use

# library(vembedr) 
# library(blogdown)

# library(crosstalk)
# library(forcats)
# library(glue)
# library(htmltools)
#library(RcppRoll) 

#library(sparkline)


## Required data

standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")

## Intermittent use

#managers<- readRDS("data/soccer/managers.rds")
#goals<- readRDS("data/soccer/goals.rds")
playerGame<- readRDS("data/soccer/playerGame.rds")
#playerClub<- readRDS("data/soccer/playerClub.rds")
#teamGames<- readRDS("data/soccer/teamGames.rds")
#teamCodes<- readRDS("data/soccer/teamCodes.rds")
# assists<- readRDS("data/soccer/assists.rds")
#scoreLines<- readRDS("data/soccer/scoreLines.rds") #








```




## Match of the Day



```{r topGame}

embed_youtube("gd4WjNSXBrg", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```


***

## The goals have gone

A [recent article in the Guardian](https://www.theguardian.com/football/blog/2018/jan/24/premier-league-swansea-liverpool-champions-league) highlighted the fact that many teams are not exceeding a goal a game this season

```{r GF_2017}

standings %>% 
   filter(tmYrGameOrder==24&season=="2017/18") %>% 
   arrange(desc(cumGF)) %>% 
   mutate(gpg=round(cumGF/tmYrGameOrder,2)) %>% 
   select(team,gpg) %>%
   plot_ly(x=~gpg,y=~fct_reorder(team,gpg),color=~team) %>% 
   add_bars() %>% 
   hide_legend() %>% 
   layout(title="Goals per Game after 24 matches",
          xaxis=list(title="Goals For"),
          yaxis=list(title=""),
          margin=list(l=100,pad=5)) %>%  config(displayModeBar = F,showLink = F)

```

In fact, over half the teams are averaging less than 1.1 per game
How does this compare, historically, at this stage of the season

```{r GF_by season}

standings %>% 
   group_by(season) %>% 
   filter(tmYrGameOrder==24&cumGF<25) %>% 
   tally() %>% 
   plot_ly(x=~season,y=~n,color=~season) %>% 
   add_bars()  %>% 
   hide_legend() %>% 
   layout(title="Teams Averaging a Goal per Game or less after 24 Matches",
          xaxis=list(title=""),
          yaxis=list(title=""),
          margin=list(l=100,pad=5)) %>%  config(displayModeBar = F,showLink = F)        
          

```
 
So, definitely one of the highest on record

***

<p class="factoid">Anthony Martial scored in his third consecutive league game, one of 19 players to achieve the feat so far this season</p>
***

## Clean Sheets

As a corollary, fewer goals by some teams will tend to mean more clean sheets by their opposition. Here are
the shutouts achieved by goalies by season. Filter by season,team or goaly using the select box


```{r}

temp <-playerGame %>% 
   filter(POSITION=="Goalkeeper"&START==6) %>% 
   left_join(standings, by="MATCHID") %>%  #
   filter(TEAMNAME!=team) %>% 
   select(LASTNAME,gameDate.x,GF,TEAMNAME,season=season.x) 
 

temp %>% 
   group_by(LASTNAME,TEAMNAME,season) %>% 
   filter(GF==0) %>% 
   count(sort=TRUE) %>% 
  #filter(season=="2017/18") %>% 
  select(season,player=LASTNAME,team=TEAMNAME,count=n) %>%
                         DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE))

         
```

At the time of writing (after 24 rounds), Manchester United's de Gea leads the way this year, with 14. Injury withstanding, he seems sure to exceed his previous best but matching the all time record of van der Sar will take some doing   

Here are the leaders per season. Hover the bar for details

```{r}

temp %>% 
   group_by(LASTNAME,TEAMNAME,season) %>% 
   filter(GF==0) %>% 
   count(sort=TRUE) %>% 
   group_by(season) %>% 
   slice(1) %>% 
   plot_ly(x=~season,y=~n,color=~TEAMNAME) %>% 
   add_bars(hoverinfo="text",
            text=~paste0(n," ",LASTNAME,", ",TEAMNAME)) %>% 
   layout(title="Goalkeeper with most clean sheets per season",
          xaxis=list(title=""),
          yaxis=list(title=""),
          margin=list(b=80,pad=5)) %>%  config(displayModeBar = F,showLink = F) 

```
_N.B. If there are two or more leading goalies in a season, only one is highlighted_

It is interesting to see the change since the early days of the Premier League when even the top performers were keeping the opposition at bay only once every four or five games 



***
<p class="factoid">The all time leader is Petr Cech with 202. Next in line is the, somewhat unhralded, Mark Schwarzer</p>
***




## Tweet of the Week 



```{r}  
shortcode("tweet", "956596478591451136") 
```



***

## Totally Football and Guardian Weekly Review Podcasts

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6612365-gratuitous-backheels-alexis-arboreal-artistry-and-the-remarkably-unremarkable-ashley-barnes/embed/v4?eid=AQAAAFguaVqN5WQA" title="Audioboom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6612474-transfer-swap-shop-another-stung-hornet-and-lyon-tamers-football-weekly/embed/v4?eid=AQAAAJouaVr65WQA" title="Audioboom player"></iframe>

***

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2018-01-16" &gameDate<="2018-01-23") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(gameDate,team) %>% 
  select(Date=gameDate,Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=400)
```

***
<p>

```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2018-01-23") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```


## Final Factoid

<p class="factoid">Mkhitaryan made a bright start to this season, leading the league with 7 assists(including secondary) in late September, but he neither added to that figure nor scored in his final 496 league minutes for United</p>


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers







