---
title: EPL Week 25
author: ''
date: '2018-02-02'
slug: epl-week-25
categories:
  - eplWeekly
tags: []
banner: 'banners/wk25_2017.png'
description: 
images: []
menu: ''

---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE,  fig.align = 'center')
```




```{r setup}


## Required

library(tidyverse) #for data carpentry
library(plotly) # for plots
library(DT) # for tables
library(vembedr) # for videos
library(blogdown) # for shortcode(tweet)







## Intermittent use

# library(vembedr) 
# library(blogdown)

# library(crosstalk)
# library(forcats)
# library(glue)
# library(htmltools)
#library(RcppRoll) 

#library(sparkline)


## Required data

standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")

## Intermittent use

#managers<- readRDS("data/soccer/managers.rds")
#goals<- readRDS("data/soccer/goals.rds")
playerGame<- readRDS("data/soccer/playerGame.rds")
#playerClub<- readRDS("data/soccer/playerClub.rds")
#teamGames<- readRDS("data/soccer/teamGames.rds")
#teamCodes<- readRDS("data/soccer/teamCodes.rds")
# assists<- readRDS("data/soccer/assists.rds")
#scoreLines<- readRDS("data/soccer/scoreLines.rds") #



```

__For the remainder of the season, I will be travelling with a back up laptop so please excuse any shortfall in posts__

***

Spurs score early - but should have more often - against a lack-lustre Manchester United sporting their shiny new toy, Alexis Sanchez 


## Match of the Day



```{r topGame}

embed_youtube("twxvaC-gP2E", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```


***

## One that got away, not

Riyad Mahrez stayed at Leicester, in spite of a transfer request and rumours that Manchester City were willing to spend somewhere around 75 million on him

There is no doubt that he has rediscovered the form he showed in the title-winning season. At the time of writing he is contributing to over 60% of his team's goals

```{r }

tmGoals <- standings %>% 
   group_by(team) %>% 
   filter(season=="2017/18"&tmGameOrder==max(tmGameOrder)) %>% 
   select(team,cumGF)
 

 
 playerGame %>% 
   group_by(name,PLAYERID,TEAMNAME) %>% 
   filter(season=="2017/18") %>% 
   summarize(points=sum(Gls)+sum(Assists)) %>% 
   rename(team=TEAMNAME) %>% 
   left_join(tmGoals) %>% 
   filter(points>0) %>% 
   mutate(pc=round(100*points/cumGF,1)) %>% 
   arrange(desc(pc)) %>% 
   ungroup() %>% 
   select(name,team,`%`=pc) %>%
                         DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE))
    


```

_N.B. Up to 2 assists may be credited to each goal_

On publishing, the under-the-radar leader is Brighton's Pascal Gross. Albeit helped by the team's poor scoring record, he has proven a shrewd purchase




***

<p class="factoid">Chelsea's surprising capitulation to Bournemouth means that they have now conceded 943 goals in the Premier League, tied with Man. City and two less than Arsenal. Spurs lead with 1252</p>
***

## Goal machines

Both Liverpool and Manchestr City recorded 3-0 victories mid-week and are tied for 
winning games by three goal margins,at 9

City just shade Liverpool 14-13 in the number of occasions they have scored at least three goals

Here are the leaders for each season at the 25 game mark

```{r}

standings %>% 
   filter(GF>=3&tmYrGameOrder<=25) %>% 
   group_by(team,season) %>% 
   tally() %>% 
   group_by(season) %>% 
   arrange(desc(n)) %>% 
   slice(1) %>% 
  plot_ly(x=~season,y=~n,color=~team) %>% 
  add_bars() %>% 
  layout(title="Team with most 3+ goal games after 25 Rounds",
         xaxis=list(title=""),
         margin=list(b=80),
         yaxis=list(title="Count")) %>%  config(displayModeBar = F,showLink = F)
 

         
```





***




## Totally Football and Guardian Weekly Review Podcasts

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6633122-swoops-swaps-and-resurgent-swans/embed/v4?eid=AQAAALr1dFqiNmUA" title="Audioboom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6633374-spurs-sweep-manchester-united-aside-plus-deadline-day-delights-football-weekly-extra/embed/v4?eid=AQAAAFX2dFqeN2UA" title="Audioboom player"></iframe>

***

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2018-01-28" &gameDate<="2018-02-02") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(gameDate,team) %>% 
  select(Date=gameDate,Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=400)
```

***
<p>

```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2018-02-01") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```


## Final Factoid

<p class="factoid">Sixth time Man. Utd. have conceded in first minute (Everton lead with 10) and second time to Spurs. Eriksen joining Jermain Defoe's effort in 2009/10</p>


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers








