---
title: EPL week 14
author: Andrew Clark
date: '2017-11-29'
slug: epl-week-14
categories:
  - eplWeekly
tags: []
banner: 'banners/wk13_2017.png'
description: 
images: []
menu: ''
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE, fig.align = 'center')
```



```{r setup}



library(tidyverse) #for data carpentry
library(plotly) # for plots
library(DT) # for tables
library(vembedr) # for videos
library(blogdown) # for shortcode
library(glue)



# library(stringr)
# library(vembedr) 
# library(blogdown)
# library(lubridate)
# library(crosstalk)
# library(forcats)
# library(glue)
# library(htmltools)

#library(sparkline)


standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")

goals<- readRDS("data/soccer/goals.rds")
playerGame<- readRDS("data/soccer/playerGame.rds")
teamGames<- readRDS("data/soccer/teamGames.rds")
# assists<- readRDS("data/soccer/assists.rds")
#scoreLines<- readRDS("data/soccer/scoreLines.rds") #



```


## Match of the Day

Allardyce is beaming. Everton are close to top half of table and he has three long-term-injured players back soon: Boalsie, Barkley and Coleman

```{r topGame}

embed_youtube("n_vQUh5X0KQ", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```

***


## Rooney Rocket

I do not collect number of goals scored from own-half but that final goal of his hat-trick was his 28th in his EPL career - good for 7th all-time 

What is perhaps even more impressive is that after just over 900 minutes on the pitch he has already notched 7 goals
In only one season has he started the year more quickly

```{r}


plDroughtFun <- function(df) {
  expand.grid(c(df$on:df$off),df$PLAYER_MATCH)
}

## calculate minuteds played to date
rooney <- playerGame %>% 
  filter(PLAYERID=="ROONEYX") %>% 
  select(season,gameDate,mins,Gls) %>% 
  group_by(season) %>% 
  arrange(gameDate) %>% 
  mutate(cumMins=cumsum(mins),cumGls=cumsum(Gls))
  

maxMins <-rooney %>% 
  filter(season=="2017/18"&cumMins==max(cumMins)) %>% 
  select(cumMins) %>% 
  pull(cumMins)

# need to add other column as are matching on TIME
gls <- goals %>% 
  mutate(scored=1)


  games <-playerGame %>% 
    filter(PLAYERID=="ROONEYX"&mins>0) %>% 
    select(PLAYERID,name,PLAYER_MATCH,START,on,offA,gameDate,TEAMMATCHID,season) %>% 
    mutate(on=as.integer(on),off=as.integer(offA)) %>% 
    mutate(on=ifelse(is.na(on),1,on),off=ifelse(is.na(off),90,off))
  
  # create list so can apply function above with purrr
  games_list <- split(games, 1:nrow(games))
  mins <-map_df(games_list,plDroughtFun)
  
  
  # rename columns and add gameDate
  names(mins) <- c("TIME","PLAYER_MATCH") 
  mins <-mins %>% 
    left_join(games) %>% 
    select(TIME,PLAYER_MATCH,gameDate,season)
  
  goalData <- mins %>% 
    left_join(gls)  %>% 
    select(PLAYER_MATCH,TIME,gameDate,scored,season) %>% 
    arrange(gameDate,TIME,season) %>% 
    group_by(season) %>% 
    mutate(minsPlayed=row_number(),goal=ifelse(is.na(scored),0,1),cumGoals=cumsum(goal))
  
  
 df <- goalData %>% 
    filter(minsPlayed==maxMins) %>% 
     arrange(season)
 
 # df %>%
 #                         DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = FALSE, searching = FALSE,info=FALSE))
 
 theTitle <- glue("Goals scored after {maxMins} minutes of season")
 
 df %>% 
   plot_ly(x=~season,y=~cumGoals) %>% 
   add_bars() %>% 
   layout(title=theTitle,
          xaxis=list(title=""),
          yaxis=list(title="Goals to date")) %>%  config(displayModeBar = F,showLink = F)
          

```

This chart will update as the season progresses

At the time of writing, however, he had yet to record an assist


***

<p class="factoid">Raheem Sterling turns 23 on 8th September. He has 95 goals and assists (up to 2 per goal) equalling Ryan Giggs at the same age. Rooney leads with 142</p>

***

## Palace Patience

After the first seven games, Palace were pointless, but doubling the number of games has seen them reach nine points - still bottom but now only fourth favourites for the drop

They recorded their first clean sheet and first away point at arch-rivals Brighton, but are now scoreless on their league travels for a record ... minutes, almost 250 minutes longer than their previous record set in 1994 and ended by current England manager, Gareth Southgate

```{r}
tmDroughtFun <- function(df) {
  expand.grid(c(1:90),df$TEAMMATCHID)
}



# need to add other column as are matching on TIME
gls <- goals %>% 
  mutate(scored=1)


games <- teamGames %>% 
  filter(TEAMNAME=="Crystal P"&venue=="A") %>% 
  ungroup() %>% 
  select(TEAMMATCHID,gameDate)


mins <- expand.grid(c(1:90),games$TEAMMATCHID)
names(mins) <- c("TIME","TEAMMATCHID")

mins <-mins %>% 
  left_join(games) %>% 
  select(TIME,TEAMMATCHID,gameDate) %>% 
  as.tbl()
names(gls)

# create gaps between goals
goalData <- mins %>% 
  left_join(gls)  %>% 
  select(TEAMMATCHID,TIME,gameDate,scored) %>% 
  arrange(gameDate,TIME) %>% 
  mutate(minOrder=row_number(),goal=ifelse(is.na(scored),0,1)) %>% 
  filter(goal==1|minOrder==max(minOrder)) %>% # to take account of current spell
  mutate(minslag=lag(minOrder),gap=minOrder-minslag)

# account for gap at beginning of career
goalData[1,]$gap <- goalData[1,]$minOrder

goalData %>% 
  arrange(gameDate) %>% 
  mutate(order=row_number()) %>% 
  plot_ly(x=~order,y=~gap) %>% 
  add_bars()  %>% 
  layout(title="Gap between away goals",
         xaxis=list(title="Goal Order"),
         yaxis=list(title="Minutes")) %>%  config(displayModeBar = F,showLink = F)
         

  


```

Maybe vary color by season (see couple of rounds ago) have it by any club (look at all games/home alternatives)


***

<p class="factoid">Giroud now has 16 goals as a sub, accounting for 23% of all his EPL strikes. Reknowned super-sub, Solskjaer, only had 19% from the bench</p>

***



***

## Tweet of the Week

```{r}  
shortcode("tweet", "936082648295866368") 
```



***


## Totally Football and Guardian Weekly Review Podcasts



<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6527110-rejuvenated-rooney-sterling-and-salah-strike-again-and-presumptuous-pep-talks/embed/v4?eid=AQAAAABgIFqGmGMA" title="audioBoom player"></iframe>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6527311-raheem-sterling-english-managers-and-the-world-cup-draw-football-weekly-extra/embed/v4?eid=AQAAAK9fIFpPmWMA" title="audioBoom player"></iframe>

***

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2017-11-21"&gameDate<="2017-11-27") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(team) %>% 
  select(Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=250)
```



```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2017-11-27") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```


## Final Factoid

<p class="factoid">Man. Utd. have already scored precisely 4 goals in six league games and look well placed to obliterate the  single-seaon record of seven which they currenly jointly hold with Chelsea and Tottenham</p>


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers
