---
title: EPL Week 20
author: Andrew Clark
date: '2017-12-27'
slug: epl-week-20
categories:
  - eplWeekly
tags: []
banner: 'banners/wk19_2017.png'
description: 
images: []
menu: ''
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE, fig.align = 'center')
```



```{r setup}



library(tidyverse) #for data carpentry
library(plotly) # for plots
library(DT) # for tables
library(vembedr) # for videos
library(blogdown) # for shortcode(tweet)







# library(stringr)
# library(vembedr) 
# library(blogdown)
 library(lubridate)
# library(crosstalk)
# library(forcats)
# library(glue)
# library(htmltools)
library(RcppRoll) 

#library(sparkline)


standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")


managers<- readRDS("data/soccer/managers.rds")
goals<- readRDS("data/soccer/goals.rds")
playerGame<- readRDS("data/soccer/playerGame.rds")
teamGames<- readRDS("data/soccer/teamGames.rds")
# assists<- readRDS("data/soccer/assists.rds")
#scoreLines<- readRDS("data/soccer/scoreLines.rds") #








```



## Match of the Day




```{r topGame}

embed_youtube("hzsCYuzgkao", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```


***
<p>


## Pardew Bounce

Alan Pardew has a rep as someone who gives a team an instant return, extended runs good and bad and ultimate disappointment
Well stage one hasn't happened with WBA as they have yet to win a game in siz but how true is it of his four other Premier managerships


```{r}

## Take account of managers who are in situ
managers[is.na(managers$Left),"Left"] <- as.Date(Sys.Date(), origin= '1970-01-01')

managerGame <-managers %>% 
  mutate(name=paste(FirstName,Lastname)) %>% 
  group_by(ManagerID,ManagerTeam) %>% 
  inner_join(standings,by=c("TEAMNAME"="team")) %>% 
  select(Lastname,FirstName,name,ManagerID,ManagerTeam,Joined,Left,TEAMNAME,gameDate,res,GF,GA,position) %>% 
  filter(gameDate>=as.Date(Joined)&gameDate<=as.Date(Left)) %>% 
  mutate(points=ifelse(res=="Win",3,ifelse(res=="Draw",1,0))) %>% 
  ungroup()



managerGame %>% 
   arrange(gameDate) %>% 
   group_by(ManagerID,name,TEAMNAME) %>% 
   mutate(cumPoints=cumsum(points),gameOrder=row_number()) %>% 
   filter(ManagerID=="PardewA") %>% 
   group_by(TEAMNAME) %>% 
   plot_ly(x=~gameOrder,y=~cumPoints,color=~TEAMNAME, colors="Accent") %>% 
  add_lines(line = list(shape = "hv")) %>% 
  layout(title="Cumulative Points by Games Played",
         xaxis=list(title='Points'),
         yaxis=list(title="Games Managed")) %>%
      config(displayModeBar = F,showLink = F)
   
    

```

For the three teams he has managed to 50 games+, there is only a 5 point span at the half-century mark

 His performaces after 6 games has been, in chronological order
 
       * West Ham United (11)
       * Charlton(8)
       * Newcastle United(10)
       * Crystal Palace(10)
       * WBA(3)
       
  So, by some margin, his worst start. The six game span has only yielded two goals
  
  Indeed, over the last 17 games, he has only registered one victory
  
```{r}
df<- managerGame %>% 
  filter(ManagerID=="PardewA") %>% 
   arrange(gameDate) %>% 
   mutate(cat=ifelse(res=="Win",1,0)) %>% 
   #group_by(ManagerID)  %>% 
   mutate(tot_sum = roll_sum(cat, 17, align = "right", fill = NA))

df %>% 
  filter(!is.na(tot_sum)) %>% 
  mutate(order=row_number()) %>% 
  plot_ly(x = ~ order,y = ~ tot_sum) %>% 
  add_markers(color= ~ TEAMNAME, colors="Accent",hoverinfo="text",symbol = I(16),
              text=~paste0(GF,"-",GA)) %>% 
  add_lines(color=I("lightgrey"),hoverinfo="text",
              text=~paste0(GF,"-",GA), showlegend=FALSE) %>% 
  layout(title="Alan Pardew - Wins in 17 game windows<br>Hover for Score",
         xaxis=list(title=""),
         yaxis=list(title="Wins")) %>%  config(displayModeBar = F,showLink = F)
         
  


```
  
 
 

***
<p class="factoid"></p>
***




```{r}
 

```




***



## 

```{r}

```





***

## Tweet of the Week


```{r}  
shortcode("tweet", "944971910609113088") 
```

This got a pretty good response with some Fulham fans aware that Clint Dempsey notched exactly 50 goals for his club

***

## Totally Football and Guardian Weekly Review Podcasts

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6564111-bristol-city-hit-the-big-time-clasico-chat-tom-ince-v-saido-berahino-and-the-doucoure-imperative/embed/v4?eid=AQAAAO7eQ1oPKWQA" title="audioBoom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6567565-christmas-carrolls-and-sid-s-ham-garage-football-weekly/embed/v4?eid=AQAAAMTbP1qNNmQA" title="audioBoom player"></iframe>

***

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2017-12-24"&gameDate<="2017-12-29") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(team) %>% 
  select(Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=250)
```

***
<p>

```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2017-12-29") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```


## Final Factoid

<p class="factoid"></p>


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers





