---
title: EPL week 11
author: Andrew Clark
date: '2017-11-06'
slug: epl2018-wk11
categories:
  - eplWeekly
tags: []
banner: 'banners/wk11_2017.png'
description: 
images: []
menu: ''
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE, fig.align = 'center')
```



```{r setup}



library(tidyverse)
library(plotly)
library(stringr)
library(vembedr) 
library(blogdown)
library(lubridate)
library(crosstalk)
library(forcats)
library(glue)
library(htmltools)

#library(sparkline)


standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")
goals<- readRDS("data/soccer/goals.rds")
playerGame<- readRDS("data/soccer/playerGame.rds")
teamGames<- readRDS("data/soccer/teamGames.rds")
assists<- readRDS("data/soccer/assists.rds")
scoreLines<- readRDS("data/soccer/scoreLines.rds") #

## for use in assister
scorer <- goals %>% 
  left_join(playerGame) %>% 
  select(scorer=name,scorerID=PLAYERID,PLAYER_MATCH_GOAL)

assister <- assists %>% 
  left_join(playerGame) %>% 
  select(name,PLAYERID,PLAYER_MATCH_GOAL,gameDate,Opponents,season)

```


## Match of the Day

Mourinho still can't buy a result away against big clubs

```{r topGame}

embed_youtube("QQ-cKUlOHHk", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```

***

## The Spanish connection

Morata connected up with Azpilicueta's cross for Chelsea's winner against Manchester United and he has now assisted on five of the striker's seven league goals this season, only topped by six of the 52 goals Diego Costa scored for the club


```{r example}
player <- "AZPILIC"

## sum for player 
scorerOrder <- assister %>% 
  filter(PLAYERID==player) %>% 
  left_join(scorer) %>% 
  group_by(scorer) %>% 
  tally() %>% 
  arrange(desc(n))


playerName<- assister %>% 
  filter(PLAYERID==player) %>% 
  head(1) %>% 
  pull(name)



assister %>% 
  filter(PLAYERID==player) %>% 
  arrange(gameDate) %>% 
  mutate(order=row_number()) %>% 
  left_join(scorer) %>% # join individ
  left_join(scorerOrder) %>% #join sum
  plot_ly(x=~order,y=~fct_reorder(scorer, n),
          hoverInfo="text",
          text=~paste0(gameDate,
                       "<br>v ",Opponents)) %>% 
  add_markers(color=~season, size=I(8)) %>% 
  layout(margin=list(l=120),
         title= glue("{playerName}'s Assists by Scorer"),
         xaxis=list(title="Goal Order"),
         yaxis=list(title="")) %>%  config(displayModeBar = F,showLink = F)



```

See the [premiersoccerstats.com](https://mytinyshinys.shinyapps.io/premierLeague/) for any player's chart



***

<p class="factoid">In spite of helping create more than 50 goals at Manchester City, de Bruyne has not assisted on any of the 7 David Silva has scored during that period</p>

***

## Comebacks

For a team in crisis, Everton did well to come back from a two goal deficit to win against Watford - even if it take an injury-time penalty miss by Cleverly to gain the points

Overall, once a deficit of two goals occurs, only 1.7% of teams recover to win the game - and there is a 93% chance of losing the game
This was Everton's third victory from 211 two goal deficits they have endured in the Premier league

Enter any team to view the outcomes of all their two goal deficits








```{r two_down}


data <-scoreLines %>% 
  group_by(TEAMMATCHID) %>% 
  mutate(
    twoDown=ifelse(diff<=-2,1,0)) %>% 
  filter(twoDown==1) %>% 
  select(TEAMMATCHID,MATCHID) %>% 
  unique()  %>% 
  left_join(teamGames) %>% 
  left_join(standings) %>% 
  select(team,GF,GA,gameDate,OppTeam,res,venue) %>% 
  mutate(GD=GF-GA) %>% 
  arrange(gameDate) %>% 
  group_by(team)

sd <- SharedData$new(data)

fs <- filter_select(
id = "team",
label = "Select Team",
sharedData = sd,
group =  ~ team,
allLevels = FALSE,
multiple = FALSE
)

 ## this is needed as crosstalk does not work nicely with bootstrap, apparently
fs_nobootstrap <- fs

attr(fs_nobootstrap, "html_dependencies") <- Filter(
  function(dep) {dep$name != "bootstrap"},
  attr(fs_nobootstrap, "html_dependencies")
)



chart <- sd %>%
   plot_ly(x=~gameDate,y=~GD) %>% 
  add_markers(color=~res,hoverinfo="text",
              text=~paste0("v ",OppTeam,"<br>",
                           GF,"-",GA)) %>% 
   layout(xaxis=list(title=""),
          yaxis=list(title="Goal Difference")) %>%  config(displayModeBar = F,showLink = F)


 tagList(
  fs_nobootstrap,
  br(),
  chart
)



```



***

## Tweet of the Week

```{r}
shortcode("tweet", "927993920084201473")
```



***


## Totally Football and Guardian Weekly Review Podcasts



<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6461984-city-sensational-again-bilic-sacked-at-last-and-seagulls-soar-of-course/embed/v4?eid=AQAAAEj1AFogmmIA" title="audioBoom player"></iframe>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6462724-slaven-bilic-gone-while-manchester-city-cruise-on-football-weekly/embed/v4?eid=AQAAACT1AFoEnWIA" title="audioBoom player"></iframe>

***

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2017-11-01"&gameDate<="2017-11-08") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(team) %>% 
  select(Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=250)
```



```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2017-11-06") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```




## Final Factoid

<p class="factoid">Crystal Palace have led in only one of the first 11 game. Even in the worst ever season - when they totalled just 11 points - Derby County led in nine of the 38</p>


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers

