---
title: EPL Week 8
author: Andrew Clark
date: '2017-10-17'
slug: epl2018-wk8
categories:
  - eplWeekly
tags: []
banner: 'banners/wk8_2017.png'
description: ''
images: []
menu: ''
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE, fig.align = 'center')
```



```{r setup}



library(tidyverse)
library(plotly)
library(stringr)
library(vembedr) 
 library(blogdown)
# library(crosstalk)
# 
# library(htmltools)

managers<- readRDS("data/soccer/managers.rds")
standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")

## might want to build earlier
managers[is.na(managers$Left),"Left"] <- as.Date(Sys.Date(), origin= '1970-01-01') #

managerGame <-managers %>% 
  mutate(name=paste(FirstName,Lastname)) %>% 
  group_by(ManagerID,ManagerTeam) %>% 
  inner_join(standings,by=c("TEAMNAME"="team")) %>% 
  select(Lastname,FirstName,name,ManagerID,ManagerTeam,Joined,Left,TEAMNAME,gameDate,res,GF,GA,position) %>% 
  filter(gameDate>=as.Date(Joined)&gameDate<=as.Date(Left)) %>% 
  mutate(points=ifelse(res=="Win",3,ifelse(res=="Draw",1,0))) %>% 
  ungroup()

# ## manager selection
# select <-managerGame %>% 
#   select(name,ManagerID) %>% 
#   arrange(name) %>% 
#   unique()
# 
# managerChoice <- select$ManagerID
# names(managerChoice) <- select$name

```


## Game of the week Crystal Palace 2 - 2 Chelsea

Champions lose to previously goalless, pointless London rivals


```{r topGame}

embed_youtube("MuLTnBYuEWs", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```

<!-- <div style="width: 100%; height: 0px; position: relative; padding-bottom: 56.250%;"><iframe src="https://www.youtube.com/watch?v=MuLTnBYuEWs" frameborder="0"  allowfullscreen style="width: 100%; height: 100%; position: absolute;"></iframe></div> -->


<p class="factoid">Julian Speroni and Gareth McAuley made their first appearances of the season becoming the oldest player and oldest outfielder, respectively, to appear in this campaign</p>



***

## Mo's MO

It is a well known that Jose Mourinho's modus operandi in away games against top sides is to play very conseravtively. It was not that succesful last season and although he got a point at Liverpool, the latter's poor defence was probably there for the taking against a team that had averaged three goals per game after seven rounds

Jamie Redknapp mentioned that you don't expect many 3-2 victories from a Mourinho team.     

 In fact, there have been six

```{r}
shortcode("tweet", "919708190715002881")
```

Well behind Sir Alex Fergsuon against whom all managers are measured and who has a marginally better overall win percentage (53-51)

Let's look at a heatmaps of their results. Hover squares for totals

```{r}


temp <- managerGame %>%
  filter(ManagerID=="MourinhoJ") %>%
   mutate(combo=paste0(GF,GA)) %>%
  group_by(combo) %>%
  tally()


allCombos <- expand.grid(
  data.frame(GF=0:9,GA=0:9)
) %>%
  mutate(combo=paste0(GF,GA)) #still a df with 100vals

test <- allCombos %>%
  left_join(temp) %>% 
  select(GF,GA,count=n)

# need to transform
Games <- t(matrix(test$count, nrow = 10, ncol = 10, byrow = TRUE,
                  dimnames = list(unique(test$GF),
                                  unique(test$GA))))


plot_ly(x = unique(test$GF), y = unique(test$GF), z = Games, key = Games, hoverinfo="z",
        colorscale='YIOrRd', reversescale=T,
        type = "heatmap") %>%
  layout(title="Mourinho Results by Goal Score",
                xaxis = list(title = "Goals Against"), 
         yaxis = list(title = "Goals For")) %>%  config(displayModeBar = F,showLink = F)
         


```


So his 23rd goalless stalemate. 2-0 victories (40) is narrowly the most common outcome



```{r}


temp <- managerGame %>%
  filter(ManagerID=="FergusonA") %>%
   mutate(combo=paste0(GF,GA)) %>%
  group_by(combo) %>%
  tally()


allCombos <- expand.grid(
  data.frame(GF=0:9,GA=0:9)
) %>%
  mutate(combo=paste0(GF,GA)) #still a df with 100vals

test <- allCombos %>%
  left_join(temp) %>% 
  select(GF,GA,count=n)

# need to transform
Games <- t(matrix(test$count, nrow = 10, ncol = 10, byrow = TRUE,
                  dimnames = list(unique(test$GF),
                                  unique(test$GA))))

plot_ly(x = unique(test$GF), y = unique(test$GF), z = Games, key = Games, hoverinfo="z",
        colorscale='YIOrRd', reversescale=T,
        type = "heatmap") %>%
  layout(title="Ferguson Results by Goal Score",
    xaxis = list(title = "Goals Against"), 
         yaxis = list(title = "Goals For")) %>%  config(displayModeBar = F,showLink = F)
         



```

Interestingly, 1-0 victories is the only result that topped the century mark but 0-0 draws were relatively less commonplace that with Mourinho 

***

## City on the Rampage

Meanwhile, United's cross-city rivals continue apace with only their second-ever sevon goal haul in the Premier League (7-0 v Norwich 2103/4) and now boast the most ever goals scored after eight rounds

```{r}


prev <- standings %>% 
  filter(tmYrGameOrder<=8&season<"2017/18") %>% 
  select(team,tmYrGameOrder,season,cumGF) %>% 
  arrange(season,tmYrGameOrder) %>% 
  group_by(season,team)

current <- standings %>% 
  filter(tmYrGameOrder<=8&season=="2017/18") %>% 
  select(team,tmYrGameOrder,season,cumGF) %>% 
  arrange(season,tmYrGameOrder) %>% 
  group_by(season,team)

manC <- standings %>% 
  filter(tmYrGameOrder<=8&season=="2017/18"&team=="Man. City") %>% 
  select(team,tmYrGameOrder,season,cumGF) %>% 
  arrange(season,tmYrGameOrder) %>% 
  group_by(season,team)

prev %>%        
plot_ly(x=~tmYrGameOrder,y=~cumGF,
        hoverinfo="text",
        text=~paste0(team,"<br>",season,"<br>",
                     cumGF," goals")) %>% 
  add_lines(color=I("lightgrey"), name="Prev Years") %>% 
  add_lines(data=current,x=~tmYrGameOrder,y=~cumGF,color=I("red"),name="2017/18") %>% 
  add_lines(data=manC,x=~tmYrGameOrder,y=~cumGF,color=I("blue"),name="Man City 17/18") %>% 
  layout(title="Goals for by team by Year",
         xaxis=list(title="Games played"),
         yaxis=list(title="Cumulative Goals")) %>%  config(displayModeBar = F,showLink = F)
         

```

Everton (2005/06) now replace Palace as the lowest scorers at this stage of the season. The Toffees had finished 4th the previous campaign


***

<p class="factoid">After 8 League games, City have already scored as many goals as 16 teams have managed after 38 matches in previous EPL campaigns - including their team of 2006/07</p>

***

## Arsenal conceding

Courtesy of a 90th minute strike by Watford's Cleverley, Arsenal(26) now only have Crystal Palace(29) ahead of them in terms of conceding the most EPL goals in 2017

```{r}
standings %>% 
  filter(venue=="A") %>% 
  select(gameDate,GA,team) %>% 
  mutate(year=str_sub(gameDate,1,4)) %>% 
  group_by(year,team) %>% 
  summarise(tot=sum(GA)) %>% 
  arrange(desc(tot)) %>% 
  filter(team=="Arsenal") %>% 
  plot_ly(x=~year,y=~tot,color=I("red")) %>% 
  layout(title="Arsenal goals conceded by calendar year",
         xaxis=list(title=""),
         yaxis=list(title="Goals conceded")) %>%  config(displayModeBar = F,showLink = F)
         
```

1992 and 2017(at the time of writing) are part years

Up to and including 2007, Arsenal conceded 20 away goals in only two years. Since then, they have only dipped below this level once.       
2011 included an 8-2 drubbing at Old Trafford

***

## Totally Football and Guardian Weekly Review Podcasts




<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6398477-stoke-enter-city-s-seventh-circle-of-swell/embed/v4?eid=AQAAAFMy7lkNomEA" title="audioBoom player"></iframe>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6398652-from-zaha-to-mourinho-the-premier-league-happiness-index-football-weekly/embed/v4?eid=AQAAANsx7lm8omEA" title="audioBoom player"></iframe>


## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2017-10-02"&gameDate<="2017-10-17") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(team) %>% 
  select(Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=250)
```



```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2017-10-17") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```




## Final Factoid

<p class="factoid">de Bruyne was withdrawn after 66 minutes with 4 assists - one shy of the EPL record</p>


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers