---
title: EPL Week 2
author: Andrew Clark
date: '2017-08-23'
slug: epl2018-wk2
categories:
  - eplWeekly
tags: []
banner: banners/wk2_2017.png
description: ''
images: []
menu: ''
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE)
```



```{r setup}



library(tidyverse)
library(forcats)
library(DT)
library(plotly)
library(vembedr)
library(lubridate)
library(stringr)
library(VennDiagram)
library(crosstalk)
library(htmltools)
library(explodingboxplotR)


playerGame <- readRDS("data/soccer/playerGame.rds")
 teamGames <- readRDS("data/soccer/teamGames.rds")
 standings <- readRDS("data/soccer/standings.rds")
 hth <- readRDS("data/soccer/hth.rds")

 ## For use in select boxes
 pgMini <- playerGame %>%  ## so wil only show those that have made an appearance - but that is prob ok
  select(PLAYERID,name) %>% 
  unique() %>% 
   arrange(name)
 
 playerChoice <- pgMini$PLAYERID
 names(playerChoice) <-pgMini$name
 
```

# Round 2

Not so many shocks in Round 2.    

 Three teams remain with 100% records. Manchester United have looked the most impressive and it is not until week 8 that they meet any of the five teams that finished above them last year.
 
 Huddersfield manager David Wagner will probably just be grateful to have six points in the bag so soon and Tony Pulis will be looking to put his feet up even earlier than usual as his WBA teams tend to do once they reach the 40 point safety level

## Game of the Week (video may not be available)

First matchup between serious title candidates and the Spurs v Chelsea game did not diappoint. Two Dele Alli inspired free kicks (one for each side) left the game at 1-1 before Lloris's gaffe provided the visitors with a win

```{r topGame}

embed_youtube("ALDZLJYZgdc", width = 420, height = 315, frameborder = 0,
  allowfullscreen = TRUE, query = NULL) 


```


## Kane Cant

Well he had eight shots but yet again no August goals for Harry Kane 

  Here are how his goals have been distributed by month



```{r kane_august}


data <- playerGame %>% 
  filter(PLAYERID=="KANEH"&mins>0) %>% 
  mutate(month=months(gameDate),monthOrder=month(gameDate)) %>% 
  select(gameDate,mins,Gls,month,monthOrder,name) %>% 
  group_by(month,monthOrder) %>% 
  summarise(apps=n(),mins=sum(mins),Gls=sum(Gls))

data %>% 
  plot_ly(x=~fct_reorder(month, monthOrder),y=~Gls,
          hoverinfo="text",
          text=~paste0("Mins: ",mins,"<br>Goals: ",Gls)) %>% 
  add_markers(size=~mins) %>% 
  layout(xaxis=list(title=""),
         yaxis=list(title="EPL Goals"),
         title="Premier League Goals by Month", # ideally get name in there
         margin=list(b=60)
         
         ) %>%  config(displayModeBar = F,showLink = F)

```

He has been hit by a fair amount of injury at the beginning of some seasons and has not really get going until November in the past

 I have extended this to cover any player

```{r}

## crosstalk version

data <- playerGame %>% 
  filter(mins>0) %>% 
  mutate(month=months(gameDate),monthOrder=month(gameDate)) %>% 
  select(gameDate,mins,Gls,month,monthOrder,name,PLAYERID) %>% 
  group_by(month,monthOrder,name,PLAYERID) %>% 
  summarise(apps=n(),mins=sum(mins),Gls=sum(Gls))

sd <- SharedData$new(data)

fs <- filter_select(
id = "player",
label = "Select or Type in Player",
sharedData = sd,
group =  ~ name,
allLevels = FALSE,
multiple = FALSE
)


## this is needed as crosstalk does not work nicely with bootstrap, apparently
fs_nobootstrap <- fs

attr(fs_nobootstrap, "html_dependencies") <- Filter(
  function(dep) {dep$name != "bootstrap"},
  attr(fs_nobootstrap, "html_dependencies")
)



monthChart <- sd %>% 
  plot_ly(x=~fct_reorder(month, monthOrder),y=~Gls,
          hoverinfo="text",
          text=~paste0("Mins: ",mins,"<br>Goals: ",Gls)) %>% 
  add_markers(size=~mins) %>% 
  layout(xaxis=list(title=""),
         yaxis=list(title="EPL Goals"),
         title="Premier League Goals by Month", # ideally get name in there
         margin=list(b=60)
         
         ) %>%  config(displayModeBar = F,showLink = F)






  tagList(
  fs_nobootstrap,
  br(),
  monthChart
)


```

Check out Rooney. He tails of at the end of the season - though Manchester United might have already taken care of business by May


<p class="factoid">WBA's Robson-Kanu joins David Lee, Klasnic and Zamora as the four subs(at the time of writing) who have scored and then seen red in Premier League history</p>

```{r sub_goal_red}

grid.newpage();

# A more complicated diagram Demonstrating external area labels
venn.plot <- draw.pairwise.venn(
  area1 = 2127,
  area2 = 62,
  cross.area = 4,
  category = c("Scored", "Dismissed"),
  fill = c("blue", "red"),
  lty = "blank",
  cex = 2,
  cat.cex = 2,
  cat.pos = c(300, 105),
  cat.dist = 0.09,
  cat.just = list(c(-1, -1), c(1, 1)),
  ext.pos = 30,
  ext.dist = -0.05,
  ext.length = 0.85,
  ext.line.lwd = 2,
  ext.line.lty = "dashed"
);
```




## Attendances

Bournemouth's small ground has held the average attendances down a bit since their promotion but the restoration of Newcastle and, more significantly, the move of Spurs to Wembley suggests records could be set this season


```{r crowds}

avCrowds <-teamGames %>%
  ungroup() %>%
  filter(venue=="H") %>%
  mutate(year=str_sub(season,1,4)) %>%
  group_by(year,TEAMNAME,season) %>%
  summarize(avCrowd=round(mean(CROWD),0)) %>%
  ungroup() %>% 
  arrange(season)



#avCrowds$year <-  factor(country.df$COUNTRY, levels = country.df$COUNTRY[order(country.df$n,country.df$COUNTRY)])

exploding_boxplot(
 avCrowds,
  y = "avCrowd",
  group = "year",
  #color = "spray", # looks like has to be a range of colors if you group- which you need to to sep data
  label = "TEAMNAME", # could be another column name but not combo
  iqr = 2,
  margin = list(bottom = 50, left = 70, top = 20, right = 20),
  xlab = "",
  ylab = "Av Attendance"
)


```

Average crowds are approaching 38,000, more than 75% higher than the inaugural season 

 Click on any box to view individual club averages by hovering the points
 


## Aguero's bogey team

With his brief sojurn out of the team at the tail end of last season, the wealth of attacking at Manchester City's disposal and the continuing rumours of possible purchase of Alexis Sanchez, it was probably not the best time for Aguero to be facing an Everton team against whom he has struggled, relatively speaking, to score.

Once again he came up empty-handed

```{r}

#this works - but does not cover zero


played<- playerGame %>% 
  filter(PLAYERID=="AGUEROS"&mins>0) %>% 
  arrange(desc(gameDate)) %>% 
  group_by(Opponents,name,PLAYERID) %>% 
  tally()  #31

playerGame %>% 
  filter(PLAYERID=="AGUEROS"&mins>0) %>% 
  arrange(desc(gameDate)) %>% 
  group_by(Opponents) %>% 
  mutate(last=row_number()) %>% 
  filter(Gls>0) %>% 
  select(Opponents,gameDate,last,Gls) %>% 
  slice(1) %>% # latest match scired in
 right_join(played) %>% 
  arrange(desc(last)) %>% 
  select(Opponents,Date=gameDate,`last scored`=last,`All Games`=n) %>% 
 DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE))



 

 

```

Next up is Bournemouth, against whom he scored two games ago

```{r}
# library(bsselectR)
# 
# bsselect(playerChoice, type = "text")

```


```{r}

# library(crosstalk)
# 
# 
# 
# 
# played<- playerGame %>% 
#   filter(mins>0) %>% 
#   group_by(Opponents,name,PLAYERID) %>% 
#   tally()  #31
# 
# allPlayers <-playerGame %>% 
#   filter(mins>0) %>% 
#   arrange(desc(gameDate)) %>% 
#   group_by(Opponents,PLAYERID) %>% 
#   mutate(last=row_number()) %>% 
#   filter(Gls>0) %>% 
#   select(Opponents,gameDate,last,Gls) %>% 
#   slice(1) %>% # latest match scired in
#  right_join(played) %>% #68032 - does not take long
#  arrange(desc(last)) 
# 
# 
# sd <- SharedData$new(allPlayers)
# 
# fs <- filter_select(
# id = "player",
# label = "Select or Type in Player",
# sharedData = sd,
# group =  ~ name,
# allLevels = FALSE,
# multiple = FALSE
# )
# 
# 
# ## this is needed as crosstalk does not work nicely with bootstrap, apparently
# fs_nobootstrap <- fs
# 
# attr(fs_nobootstrap, "html_dependencies") <- Filter(
#   function(dep) {dep$name != "bootstrap"},
#   attr(fs_nobootstrap, "html_dependencies")
# )
# 
# 
# 
# 
#  #  produce table
#  # myTable  <-  sd %>%
#  #     #arrange(desc(last)) %>%
#  #   #select(Opponents,Date=gameDate,`last scored`=last,`All Games`=n) %>%
#  #  DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE))
# 
#  # Error in DT::datatable(., class = "compact stripe hover row-border order-column",  : 
#  #  'data' must be 2-dimensional (e.g. data frame or matrix)
#  
#  # works - though seems to show all teams
#  # myChart  <-  sd %>% 
#  #    plot_ly(x=~Opponents,y=~last) %>% 
#  #    add_markers()
#  # 
# 
# myTable  <-  sd %>%
#   DT::datatable()  #Error in DT::datatable(.) : 
#   #'data' must be 2-dimensional (e.g. data frame or matrix)
# 
#   tagList(
#   fs_nobootstrap,
#   br(),
#    myTable
# )
# 


```

---

## Totally Football and Guardian Weekly Review Podcasts




<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6225033-man-united-looking-menacing-ill-advised-celebrations-and-west-brom-s-axis-of-evil/embed/v4?eid=AQAAAPyVm1mJ_F4A" title="audioBoom player"></iframe>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6225029-is-the-wembley-curse-fake-news-football-weekly/embed/v4?eid=AQAAAD2Wm1mF_F4A" title="audioBoom player"></iframe>

---

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2017-08-15"&gameDate<="2017-08-21") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(team) %>% 
  select(Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=250)
```



```{r standings}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2017-08-21") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```




## Final Factoid

<p class="factoid">The +8 goal difference for Man Utd after two games is their best ever in the Premier League and trails only the 2010/11 Chelsea side that inflicted 6-0 lickings to WBA and Wigan to start that season</p> 


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers