---
title: EPL Week 1
author: Andrew Clark
date: '2017-08-15'
slug: epl2018-wk1
categories:
  - eplWeekly
tags: []
banner: banners/wk1_2017.PNG
description: ''
images: []
menu: ''
---


```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE)
```



```{r setup}


# library(DBI)
# library(odbc)
# library(dplyr)
# library(dbplyr)
library(tidyverse)
library(forcats)
library(DT)
library(plotly)
library(vembedr)
# library(purrr)
# library(engsoccerdata)

playerGame <- readRDS("data/soccer/playerGame.rds")
teamGames <- readRDS("data/soccer/teamGames.rds")
standings <- readRDS("data/soccer/standings.rds")
hth <- readRDS("data/soccer/hth.rds")

```

# Round 1

Well we're back with an exciting start to the new season. Manchester United turned in their best performance in  years and Huddersfield had a dream return to the top tier. Some things remain the same though with Dele Alli conning Shelvey into a dismissal and Liverpool conceding from corners

## Game of the Week

Arsenal and Leicester swapping blows with the Gunners emerging victorious

```{r topGame}

embed_youtube("tXq8TlZupEY", width = 420, height = 315, frameborder = 0,
  allowfullscreen = TRUE, query = NULL) 


```






## Foreign Players

One fairly constant theme has been the downturn in English player over the years. 



```{r english_starters}


games <- teamGames %>% 
  ungroup() %>% 
  filter(tmYrGameOrder==1) %>% 
  pull(MATCHID)

 playerGame %>% 
  filter(MATCHID %in% games&START>0) %>% 
  group_by(season,COUNTRY) %>% 
  summarise(tot=n()) %>% 
  group_by(season) %>% 
  mutate(pc=round(100*tot/sum(tot),1)) %>% 
  filter(COUNTRY=="England") %>% 
  ungroup() %>% 
  plot_ly(x=~season,y=~pc) %>% 
  add_lines() %>% 
  layout(margin=list(b=70),
         title="% of Starters born in England - EPL Round 1",
    xaxis=list(title=""),
    yaxis=list(title="% ")
  ) %>%  config(displayModeBar = F,showLink = F)

```

This appears to have stabilized over the past couple of seasons

---

Check out how each of the teams started the season


```{r country_table}

# 1st weeek tis year
playerGame %>% 
  filter(MATCHID %in% games&START>0&season=="2017/18") %>% 
  #mutate(place=glue({city},{COUNTRY})) %>% 
  mutate(place=paste0(city,", ",COUNTRY)) %>% 
  select(team=TEAMNAME,player=LASTNAME,place) %>% 
  arrange(team,player)%>%
                         DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE,lengthMenu=c(11,11)))


# playerGame %>% 
#   filter(MATCHID %in% games&START>0&season=="2017/18") %>% 
#   group_by(TEAMNAME,COUNTRY) %>% 
#   tally() %>% 
#   ungroup() %>% 
#   select(TEAMNAME) %>% 
#   group_by(TEAMNAME) %>% 
#   tally()

# 
# playerGame %>% 
#   filter(TEAMNAME=="Arsenal"&START>0) %>% 
#   group_by(gameDate,COUNTRY) %>% 
#   tally() %>% 
#   arrange(desc(n)) %>% 
#   filter(COUNTRY=="France")



```

New-signing, Lacazette, wass the only Frenchman in Arsenal's starting line-up (It has been as high as seven). 

 Stoke boasted ten different nationalities
 
---

Here is the breakdown across the EPL

```{r fig.height=10}

country.df <- playerGame %>% 
  filter(MATCHID %in% games&START>0&season=="2017/18") %>% 
  group_by(COUNTRY) %>% 
  tally() 

#use forcats - works even though character - then prder on reverse alphabetically - ideally opposite of that
country.df  %>%
  plot_ly(x=~n,y=~fct_reorder(COUNTRY, n),height=1000) %>%
  layout(margin=list(l=100,pad=5),
  yaxis=list(title=""),
  xaxis=list(title="")) %>%
   config(displayModeBar = F,showLink = F)
  

```


No clear candidate for second place but newly-promoted Huddersfield coach, David Wagner, has signed a few fellow-Germans

Only two Scots, Fraser (Bournemouth) and 33-year-old, Fletcher (Stoke). The last time Scotland held the second rank, after England, in a game to lead off a season was in 1999



```{r 2nd_place}


 # playerGame %>% 
 #  filter(MATCHID %in% games&START>0) %>% 
 #  group_by(season,COUNTRY) %>% 
 #  summarise(tot=n()) %>% 
 #  group_by(season) %>% 
 #  arrange(desc(tot)) %>% 
 #  slice(2)

```


## Big Away Wins


<p class="factoid">What Birmingham never did in seven seasons and Southampton only achieved at the 186th attempt, Huddersfield accomplished at the first time of asking<br>Won their first EPL away game by at least a three goal margin<br>Surprisingly, three other teams match them, including Blackpool, who were relegated the same year</p> 



```{r, fig.height= 12}


# first games for teams that have ghad 3 goals margins
didIt <-standings %>% 
  arrange(gameDate) %>% 
  filter(venue=="A") %>% 
  group_by(team) %>% 
  mutate(awayGameOrder=row_number()) %>% 
    filter((GF-GA)>=3) %>% 
  group_by(team) %>% 
    slice(1)

# teams that have never achieved  
never <-setdiff(unique(standings$team),unique(didIt$team))

# never teams away games to date - select relevant cariables+ add ever
neverTeams <- standings %>% 
  filter(team %in% never&venue=="A") %>% 
  group_by(team) %>% 
  arrange(tmGameOrder) %>% 
  mutate(awayGameOrder=row_number()) %>% 
  arrange(desc(awayGameOrder)) %>% 
  slice(1) %>% 
  select(team,awayGameOrder) %>% 
  mutate(ever="Never")

# cf above re fields
didItTeams <- didIt %>% 
  select(team,awayGameOrder) %>% 
  mutate(ever="First")

# combine
df <- rbind(neverTeams,didItTeams)

## order team as factor for chart
df$team <-  factor(df$team, levels = df$team[order(df$awayGameOrder)])


 #chart 
df %>% 
  plot_ly(x=~awayGameOrder,y=~team,height=1000) %>% 
  add_bars(color=~ever) %>% 
  layout(title=" EPL Away Games before 3 goal victory",
        margin=list(l=120,pad=5),
         xaxis=list(title="Games"),
         yaxis=list(title="")
         
         ) %>%  config(displayModeBar = F,showLink = F)


```

It took Newcastle more than a century of matches which is a touch unexpected given that they were a high-scoring, successful team when they first made the Premier league

Southampton definitely buck the trend taking 186 games to reach the target - though they have had more success recently

```{r}

standings %>% 
    filter(venue=="A"&team=="Southampton") %>% 
  mutate(GD=GF-GA) %>% 
  group_by(GD) %>% 
  tally() %>% 
  plot_ly(x=~GD,y=~n) %>% 
  add_bars(color=I("red3")) %>% 
  layout(title="Southampton PL results by Goal Difference",
         xaxis=list(title="Goal Difference"),
         yaxis=list(title="Games Count")) %>%  config(displayModeBar = F,showLink = F)


```

## Man Utd are back

I know, only West Ham, but still far from the turgid play of recent seasons. They end by far their longest run without a four goal victory


```{r mnu_streak}

standings %>% 
  arrange(gameDate) %>% 
  group_by(team) %>% 
  filter((GF-GA)>=4&team=="Man. Utd.") %>% 
  select(gameDate,tmGameOrder) %>% 
  mutate(last=lag(tmGameOrder)) %>% 
  mutate(last=(ifelse(is.na(last),0,last)),diff=tmGameOrder-last,order=row_number()-1) %>% 
  plot_ly(x=~order,y=~diff) %>% 
  add_bars(color=I("red")) %>% 
  layout(title="Man Utd sequences between 4 goal victories",
         xaxis=list(title=""),
         yaxis=list(title="")) %>%  config(displayModeBar = F,showLink = F)

```

---

## Totally Football and Guardian Weekly Review Podcasts




<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6203313-chelsea-crash-and-burn-ley-as-huddersfield-make-their-mark/embed/v4?eid=AQAAAO9tlFmxp14A" title="audioBoom player"></iframe>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6203084-goals-petulance-and-absent-defences-the-premier-league-is-back-football-weekly/embed/v4?eid=AQAAANRslFnMpl4A" title="audioBoom player"></iframe>

---

## Results and Table

```{r results}
## results - amy have to copy and run in console
hth %>% 
  filter(season=="2017/18"&gameDate>="2017-08-09"&gameDate<="2017-08-15") %>% ## may need to put in day later?
  filter(venue=="H") %>% 
  arrange(team) %>% 
  select(Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=250)
```

---

```{r table}


# table
hth %>% 
  filter(season=="2017/18"&gameDate<="2017-08-17") %>% 
  group_by(team) %>% 
  mutate(W = ifelse(res=="Win",1,0),L = ifelse(res=="Loss",1,0),D = ifelse(res=="Draw",1,0)) %>%
  summarise(P=n(),Pts=sum(points),W=sum(W),D=sum(D),L=sum(L),GD=sum(GF)-sum(GA),GF=sum(GF)) %>% 
  arrange(desc(Pts),desc(GD),desc(GF),team) %>%
  DT::datatable(class='compact stripe hover row-border order-column',colnames = c('', 'P', 'Pts', 'W','D', 'L', 'GD','GF'),
                rownames=TRUE,options= list(paging = FALSE, searching = FALSE,info=FALSE))


```




## Final Factoid

<p class="factoid">Crystal Palace have won just nine of their most recent 33 league games. More than half these victories came in  the 7 games in which Sakho played a full ninety minutes</p> 


---



Plenty more NEW data at the [premiersoccerstats site](https://mytinyshinys.shinyapps.io/premierLeague/)

Feel free to contact me if you have any questions or work offers