---
title: geofacetting South Africa style
author: Andrew Clark
date: '2018-03-27'
slug: geofacetting-south-africa-style
categories:
  - opendata
tags:
  - geofacet
banner: ''
description: ''
heimages: []
menu: ''
---


Not sure if I have blogged about Ryan hafens has recently started what is promised to be a series of blog posts about his geofacet package. In the first post, he argues the case for why this approach is superior to other forms of choropleth. I have previously...

It is possible to create one's own map to apply yo geofacetting and luckily one has already been created for the nine provinces of South Africa - where I am currently visiting

There is also a portal Nesstar where it is possible to download lots of census and household survey data in many tpes of format including csvs

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE,  fig.align = 'center')
```

```{r setup}
library(tidyverse)


library(geofacet)


```

There is also a portal Nesstar where it is possible to download lots of census and household survey data in many tpes of format including csvs/ Along with a pdf describing the data

## 2016 Household survey

The 12MB takes a few seconds to read in but results in a data.frame with 274 columns so there
are quite a lot of variables available to plot. 

Lets summarize by province the variable, Q52Roof, which indicates for each
household the construction material for the roof


```{r}

df <- read_csv("data/GHS-2016-HOUSE_F1.csv")

df %>% 
  group_by(Prov,Q52Roof) %>% 
  tally()


```

Well that looks okay but there are a few points to address

  * The Provinces need to be named
  * The Type of Roof needs to be described
  * The Numbers need to be a percentage - Combining the small values will make chart neater
  
```{r roof_processing}

# I have already created a lookup table for provinces

provinces <- read_csv("data/sa_provinces.csv")

# The above-mentioned pdf provides the description for each of the Roof codes

roofTypes <- tibble(code=c(1:13,99),description =c(
  "Bricks",
  "Cement",
  "Corrugated",
  "Wood",
  "Plastic",
  "Cardboard",
  "Mud/Cement",
  "Wattle",
  "Tile",
  "Mud",
  "Thatch",
  "Asbestos",
  "Other",
  "Unspecified")
)

## Lets do some data-carpentry

roofs <- df %>% 
  group_by(Prov,Q52Roof) %>% 
  tally() %>% 
  left_join(provinces,by=c("Prov"="prov_id")) %>% 
  left_join(roofTypes,by=c("Q52Roof"="code")) %>% 
  select(province,description,n) %>% 
  #ungroup() %>% 
  group_by(province) %>% 
  mutate(pc=round(100*n/sum(n),1)) %>% 
  #ungroup() %>% 
         mutate(description = ifelse(pc > 3, description, "Other")) %>% 
  group_by(description,province) %>% 
  summarise(pc = sum(pc))

roofs

```
  
  
  We now have data we can plot using ggplot and geofacet
  
```{r roofs_plot}

roofs %>% 
  ggplot(aes(description,pc,fill=description)) +
  geom_col() +
  facet_geo(~ province, grid = "sa_prov_grid1") +
  labs(title = "South Africa % Roof Types by province",
    caption = "Data Source: SA 2016 Household Survey",
    y = "%",
    x="") +
   theme_bw() +
   theme(
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) 


```
  
  So Corrugated Iron/Zinc rules - associated with shacks - rules other than in Cape Town