---
title: User2017- padr package example
author: Andrew Clark
date: '2017-07-19'
slug: user2017-padr
banner: banners/padr.PNG
---

Of course, it is not the same as actually being there, but as a good fall-back the videos of the talks for the R User 2017 conference are now available on [channel 9](https://channel9.msdn.com/Events/useR-international-R-User-conferences/).    

I'll be dipping into them over the next few weeks and reporting on any I find of interest.     

Let's kick-off with the [padr package](https://github.com/EdwinTh/padr) from [Edwin Thoen](https://twitter.com/edwin_thoen). It is on CRAN

> padr is an R package that assists with preparing time series data. It provides two main functions that will quickly get the data in the format you want. When data is observed on too low a level, thicken will add a column of a higher interval to the data frame, after which the user can apply the appropriate aggregation. When there are missing records for time points where observations were absent, pad will automatically insert these records. A number of fill_ functions help to subsequently fill the missing values.



```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE)

```

Load libraries 

```{r setup}
library(tidyverse)
library(vembedr)

library(plotly)

library(padr)



```



Here is his lightning talk

```{r talk }

embed_user2017("Room-301-Lightning-Talks", width = 560, height = 315, frameborder = 0,
  allowfullscreen = TRUE) %>% 
  use_start_time("5m20s")

```

My example uses a small dataset of the season debut of players for clubs in the English Premier League over the past two campaigns

```{r}
df <- read_csv("data/soccer/seasonDebut.csv")
head(df)

```

The raw data shows dates for a given day and I want to produce a plot - a line chart like Edwin's example - but collating the data on a less granular basis.    

Let's first look at the data applying the thicken() function which places the data into a higher interval. The full range available is year, quarter, month, week, day, hour, minute, or second. in this instance, 'month' seeems the most appropriate 

As there is only one date-time field, it does not need to be specified

```{r thicken}

df_month <- df %>% 
    thicken('month') %>% 
    group_by(gameDate_month) %>% 
    summarise(month_count = n()) 

df_month %>%
                         DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = FALSE,info=FALSE))
   
df_month %>% 
    ggplot(aes(gameDate_month, month_count)) + geom_line() 


```

Not unexpectedly, most of the season debuts are in the first few weeks of each year i.e. in August. However, you can see a problem here in that the chart spans an - off-season of June and July 2016 - which the line chart does not cater for    

Enter the pad() function. The default values are again appropriate


```{r pad}

df_month <- df %>% 
    thicken('month') %>% 
    group_by(gameDate_month) %>% 
    summarise(month_count = n()) %>% 
    pad()


df_month %>%
                         DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = FALSE,info=FALSE))
   
df_month %>% 
    ggplot(aes(gameDate_month, month_count)) + geom_line() 


```

The missing months are now included (with NA values) and the chart now reflects reality


In this example, a barchart would probably be more appropriate - and would not have necessitated the pad() code. 

Here is a plotly version

```{r plotly}

df_month %>% 
  plot_ly(x=~gameDate_month, y=~month_count) %>% 
  add_bars() %>% 
  layout(title="Monthly season debuts of EPL players",
         xaxis=list(title=""),
         yaxis=list(title="Count")) %>% 
   config(displayModeBar = F,showLink = F)

```

More than 70% of players make their debuts in August. There are slight tick-ups in January and February - in line with new players acquired in the transfer window - and at the end of season, when youngsters are often introduced in relatively meaningless games

So, all-in-all a simple but useful addition to the R toolkit