---
title: What the deuce!
author: Andrew Clark
date: '2018-03-25'
slug: what-the-deuce
categories:
  - sports
  - tennis
tags: []
banner: ''
description: ''
images: []
menu: ''
---

At the recent [Satrday conference](https://capetown2018.satrdays.org/) in Cape Town, [Stephanie Kovalchik](@StatsOnTheT) gave a keynote speech
relating to her work at Tennis Australia

One of her accomplishments  is the [deuce package](https://github.com/skoval/deuce)  which is a 'collection of data sources and tools for extracting data for professional men's and women's tennis'

I thought I would take it for a spin

First load the packages required - including crosstalk and htmltools for interactivity

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE,  fig.align = 'center')
```

```{r setup}
library(tidyverse)
library(plotly)

library(deuce)

library(crosstalk)
library(htmltools)


```

There are several datasets.

## Men's ELO Ratings

For this post, I am going to commence with the ELO rankings.  Unlike the official rankings, Elo ratings give credit for who you play, not the round or tournament in which you play them. However, like any rankings, any comparison
over different time-periods should be treated with caution

The period covered is the Open era - when all tournaments had prize money and the elite professionals - like Rod Laver and Lew Hoad - rejoined the main circuit

Let's check out the men first

```{r tables}

#data("atp_players")

data("atp_elo")


head(atp_elo)

atp_elo %>% 
  count(player_name) %>% 
  plot_ly(x= ~n) %>% 
  add_histogram() %>% 
  layout(title="Distribution of Player ATP Match careers",
         xaxis=list(title="Number of Matches"),
         yaxis=list(title="Player Count")) %>% config(displayModeBar = F, showLink = F)
  


```

There are over 300,000 rows, featuring more than 4,000 players - most of whom
have records of 5 matches or list in the datset. Jimmy Connors leads with  1462 betwen 1970 and 1996

Toggle back to data.frame window to view variable details

---

### Andy Murray

I am first going to take a look at Britain's greatest tennis player during this era, Andy Murray.

I need to do some processing on the raw atp_elo data to get one ranking per tournament and process the 'round' info to be more meaningful in a chart

```{r Murray}

# get ELO rating at end of each tournament
murray <-atp_elo %>% 
  filter(player_name=="Andy Murray") %>% 
  group_by(tourney_start_date) %>% 
  slice(n())

# create new factor variable for stage at which he exits tournament

murray <- murray %>% 
  mutate(exit= case_when(
         str_sub(round,1,1) == "R"&str_sub(round,2,2) != "R" ~ str_sub(round, start = 2),
         round == "QF" ~ "Quarters",
         round == "SF" ~ "Semis",
         round == "F"&win == 0 ~ "Final",
         round == "F"&win == 1 ~ "Winner",
         TRUE                     ~ "Unknown"
         )
  ) %>% 
  mutate(exit=as.factor(exit))

# create better order for legend
revOrder <- c("Winner","Final","Semis","Quarters","16","32","64","128","Unknown")

  murray %>% 
    plot_ly(x=~ tourney_start_date, y = ~overall_elo) %>% 
    add_markers(hoverinfo="text", color =~ fct_relevel(exit, revOrder),
                 colors ="Spectral",
                text=~paste(
                  tourney_name," - ",year(tourney_start_date),
                  "<br>v ",opponent_name,
                  "<br> Round Reached: ",exit,
                   "<br>ELO: ",round(overall_elo)))  %>% 
    layout(title="ELO Rankings over time - Andy Murray",
            xaxis=list(title="Tournament Start Date"),
            yaxis =  list(title = "ELO Ranking")
           ) %>% 
     config(displayModeBar = F, showLink = F)
                  
                
  


```

Hover the data for more info - specifically relating to his final game in each tournament

You can double-click on the legend to isolate the round he reached at each tournament. Interstingly, he has lost in 10 finals since Wimbledon 2012 - all bar one to Novak Djokevic

### Other Top 100 all time

We can extend this to all players - although for speed of processing I have restricted to the top 100

```{r men_elo}


# create vector of top 100 players
# based on their max elo rating

bestEver <-atp_elo %>% 
  group_by(player_name) %>% 
  arrange(desc(overall_elo)) %>% 
  slice(1) %>% 
  arrange(desc(overall_elo)) %>% 
  ungroup() %>% ## otherwise slice does not appear to work
  slice(1:100) %>% 
  pull(player_name)


df <-atp_elo %>% 
      filter(player_name %in% bestEver) %>% 
    mutate(exit= case_when(
         str_sub(round,1,1) == "R"&str_sub(round,2,2) != "R" ~ str_sub(round, start = 2),
         round == "QF" ~ "Quarters",
         round == "SF" ~ "Semis",
         round == "F"&win == 0 ~ "Final",
         round == "F"&win == 1 ~ "Winner",
         TRUE                     ~ "Unknown"
         )
  ) %>% 
  group_by(tourney_start_date,player_name) %>% 
  slice(n()) 

# create sharedData for interactivity
 sd <- SharedData$new(df)


# select input details
 fs <- filter_select(
id = "player_name",
label = "Select Player",
sharedData = sd,
group =  ~ player_name,
allLevels = FALSE,
multiple = FALSE
)

 ## this is needed as crosstalk does not work nicely with bootstrap
fs_nobootstrap <- fs

attr(fs_nobootstrap, "html_dependencies") <- Filter(
  function(dep) {dep$name != "bootstrap"},
  attr(fs_nobootstrap, "html_dependencies")
)

#print(player_name)
print(sd$player_name)

#create chart
chart <-  sd %>% 
    plot_ly(x=~ tourney_start_date, y = ~overall_elo,height=300) %>% 
    add_markers(hoverinfo="text",  color =~ fct_relevel(exit, revOrder),
                 colors ="Spectral",
                text=~paste(
                  tourney_name," - ",year(tourney_start_date),
                  "<br>v ",opponent_name,
                  "<br> Round Reached: ",exit,
                   "<br>ELO: ",round(overall_elo)))  %>% 
    layout(title="ELO Rankings over time",
            xaxis=list(title="Tournament Start Date"),
            yaxis =  list(title = "ELO Ranking")) %>% 
     config(displayModeBar = F, showLink = F)


# display correctly
tagList(
  fs_nobootstrap,
  br(),
  chart
)




 




```


Currently, the default is to show all players - so make sure to select only one from the input selector!

## Women Rankings by Country

Let's look at another dataset - world rankings; this time concentrating on the women and in particular, whether any one country has ruled the roost

This will necessitate utilizing a couple of the other datasets. Information is available from the mid 1980's

I have averaged each player's ranking points over the year and then selected the top ten players

```{r}

data("wta_players")
data("wta_rankings")

## mean ranking points in year from  1986

wta_rank <- wta_rankings %>% 
  mutate(year=year(date)) %>% 
  group_by(player_id,year) %>%
  summarize(avPoints=mean(ranking_points,na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(year) %>% 
  arrange(desc(avPoints)) %>% 
  slice(1:10) %>% 
  filter(year>1985) %>% 
  left_join(wta_players) %>% 
  select(name,country_code,year,avPoints)

# Now collate by country by year and make a plot

wta_rank %>% 
  count(country_code,year) %>% 
  group_by(country_code) %>% 
  plot_ly(x=~year,y=~n, color=~country_code) %>% 
  add_bars() %>% 
  layout(barmode = 'stack',
         title="Distribution of Top Ten WTA Players<br> by
         Country by Year",
         xaxis=list(title=""),
         yaxis =  list(title="Count")) %>% config(displayModeBar = F, showLink = F)

```

     Double-click on the legend (which you can scroll down to see more countries) to see individual countries by year. The USA has only missed out in one season but has relied on Serena Williams (now 36) for representation in recent seasons. Russia had a good spell earlier this century
     
---

Below is a searchable, sortable table that shows individuals who have contributed to the above graph

```{r top_ten_table}

wta_rank %>% 
  arrange(desc(avPoints)) %>% 
  group_by(year) %>% 
  mutate(rank=row_number(),Points=round(avPoints,0)) %>%
  select(Player=name, Country = country_code,Year=year,Points,
         `Annual Rank`=rank) %>% 
  DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE))

```

Again, comparing across time-periods for relative ability is probably not that relevant

Great Britain, which had some Majors champions, like Wade and Jones, prior to the period under consideration did not have a single top ten player until Konta in 2017